---
title: "N uptake calculations"
author: "Zhang Zhenglin"
output:
  pdf_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
editor_options:
  markdown:
    wrap: sentence
---

# Necessary libraries

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(ggplot2)
theme_set(theme_bw())
library(emmeans)
library(multcomp)
library(PLS205)
library(lme4)
library(lmerTest)
library(multcompView)
library(car)
library(Rmisc) 
library(dplyr) #https://r4ds.had.co.nz/ (Chapter 3, Chapter 5, look at filter and select)
library(stringr) 
library(readxl)
library(writexl)
```

# Read and check data

```{r}
#read data
master = read_excel('SIF_NRate_NUptake_Base_Dataframe.xlsx', sheet = 1)

master$Sample_Weight_mg <- as.numeric(master$Sample_Weight_mg)                     

str(master)

table(master$Stage)
table(master$Treatment)
table(master$Topdress)
table(master$Blk)
table(master$Year)


#dataset looks good and complete
```

# N uptake calculations

N uptake for PI and 50% heading are determined from non-segregated plant parts.
i.e. the whole plant was ground to a powder together and is representative of the crop's N uptake.
As most of the tissue here is vegetative, we assume that ambient moisture is the same as straw (8.27%)

N uptake for maturity are determined from separated plant parts - grain and straw.
This step calculates N present in grain and straw, but do not sum them yet.
Moisture for is 8.27% for straw and 9.76% for grains.

Dry samples were measured after stabilizing in the oven for a period of time (\>24hrs) and was measured to be 1.9%.

For 2023, we directly used yield.
Grain yield was reported at 14%

For 2023, grain was harvested by combine.
A sub-sample was taken to evaluate N uptake.
Harvest index (HI) was quantified from this subsample and used to estimate straw weight.

For for information on moisture calculations: <https://edis.ifas.ufl.edu/publication/AG442>

```{r}
#Getting: g N per g of microbalanced sample (g/g)

master$N_per_g_microbalanced_sample <- ((master$SIF_N_µg)/(master$Sample_Weight_mg))/1000

#Dry sample weight is at 1.9% for 2021 and 2022 samples, 14% for 2023 samples
#We need to correct the weight for ambient moisture where the samples were measured into tin capsules

master$Ambient_Sample_Weight_g <- master$Dry_Sample_Weight_g*((100-master$Dry_sample_moisture_percent)/(100-master$Microbalance_moisture_percent))

str(master)
```

```{r}
#Total N uptake = g N per g of microbalanced sample (g/g)*total sample weight (g)/area harvested (m2) 
#then scale from g/m2 to kg/ha

master$N_total_kgha <- ((master$N_per_g_microbalanced_sample*master$Ambient_Sample_Weight_g)/1000) * (10000/master$Area_harvested_m2)

str(master)
```

# Getting the maturity summed values: summing N uptake values of straw and grain

```{r}
#separate into straw and grains for summing
maturity_straw <- master %>%  filter(Stage == ("Maturity_Straw"))
maturity_grain <- master %>%  filter(Stage == ("Maturity_Grain"))


#maturity_straw <- 
 # maturity_straw %>% 
  #filter(Year != 2021) %>%
 # mutate(Plot = substr(SampleID, 1, 3)) 

#maturity_grain <- 
 # maturity_straw %>% 
  #filter(Year != 2021) %>%
  #mutate(Plot = substr(SampleID, 1, 3)) 

as.numeric(maturity_straw$Plot) - as.numeric(maturity_grain$Plot) 

#Create dataframe for maturity total N uptake
maturity_summed <- as.data.frame(rbind(master[1:36,], master[109:156,], master[241:288,])) 
# just so we can keep the other columns
maturity_summed$Stage <- "Maturity"

#sum grain and straw into one maturity total N uptake dataframe
maturity_summed$N_total_kgha <- maturity_straw$N_total_kgha+maturity_grain$N_total_kgha

str(maturity_summed)
```

# Bind everything back again - this is such a pain in the rear end. Add in plots and Nrates

```{r}
#Well I was a clown when I started the PhD, so I named everything for excel LOL. 
#Because name formatting was different by year, I split the dataframe by year and wrangle it as needed

# extract PI and Heading N uptake values
master_maturity_summed <- rbind(master[73:108,], master[205:240,], maturity_summed)

#add in plot and treatments

master_maturity_summed_2021 <- 
  master_maturity_summed %>% 
  filter(Year == 2021) %>%
  mutate(Plot = substr(SampleID, 5, 7)) %>%
  mutate(Treatment = case_when(
  Plot %in% c("101", "102", "103", "104", "105", "106",
              "201", "202", "203", "204", "205", "206",
              "301", "302", "303", "304", "305", "306") ~ "FR",
  Plot %in% c("107", "108", "109", "110", "111", "112",
              "207", "208", "209", "210", "211", "212",
              "307", "308", "309", "310", "311", "312") ~ "CR",
  TRUE ~ "Other" # This line handles cases where plot is not listed
  )) 

master_maturity_summed_others <- 
  master_maturity_summed %>% 
  filter(Year != 2021) %>%
  mutate(Plot = substr(SampleID, 1, 3)) %>%
  mutate(Treatment = case_when(
  Plot %in% c("401", "402", "403", "404", "405", "406",
              "501", "502", "503", "504", "505", "506",
              "601", "602", "603", "604", "605", "606",
              "701", "702", "703", "704", "705", "706",
              "801", "802", "803", "804", "805", "806",
              "901", "902", "903", "904", "905", "906") ~ "FR",
  Plot %in% c("407", "408", "409", "410", "411", "412",
              "507", "508", "509", "510", "511", "512",
              "607", "608", "609", "610", "611", "612",
              "707", "708", "709", "710", "711", "712",
              "807", "808", "809", "810", "811", "812",
              "907", "908", "909", "910", "911", "912") ~ "CR",
  TRUE ~ "Other" # This line handles cases where plot is not listed
  )) 

master_maturity_summed_clean <- rbind(master_maturity_summed_2021, master_maturity_summed_others)

# add in blk, days

master_maturity_summed_clean <- master_maturity_summed_clean %>%
  mutate(Blk = substr(Plot, 1, 1))%>%
  mutate(Days = case_when(
  Stage %in% c("PI") ~ "47",
  Stage %in% c("Heading") ~ "82",
  Stage %in% c("Maturity") ~ "134",
  TRUE ~ "Other" # This line handles cases where plot is not listed
  ))

# read in Nrate reference file

#clean the data frame to have one plot match with one Nrate
Nrate_ref <- read_excel("D:/Academics/UC Davis/School Work/Linquist Lab/Data/R stats/Agronomic paper/Yield/master_yields.xlsx", sheet = 1) %>%  mutate(Plot = substr(Plot, 1, 3)) %>% mutate(Nrate_kgha = Nrate) %>%  distinct(Plot, .keep_all = TRUE)

# put in the Nrate for all the plots
master_maturity_summed_clean <- master_maturity_summed_clean %>%
  left_join(Nrate_ref %>% select(Plot, Nrate_kgha), by = "Plot") 
```

# Export dataframe

```{r}
for_export <- master_maturity_summed_clean %>% 
                         select(SampleID,
                                Plot,
                                Stage,
                                Treatment,
                                Year,
                                Topdress,
                                N_total_kgha,
                                Blk,
                                Days,
                                Nrate_kgha)

head(master_maturity_summed_clean)

write_xlsx(for_export, "N_uptake_Maturity_Summed_Clean.xlsx")
```

# Filtering for N uptake over time

```{r}
N_over_time <- master_maturity_summed_clean %>% 
               filter(Topdress == '0') %>%
               select(SampleID,
                                Plot,
                                Stage,
                                Treatment,
                                Year,
                                Topdress,
                                N_total_kgha,
                                Blk,
                                Days,
                                Nrate_kgha)

write_xlsx(N_over_time, "D:/Academics/UC Davis/School Work/Linquist Lab/Data/R stats/Agronomic paper/N uptake/N uptake over time - 2021_22/N_uptake_over_time.xlsx")
```

# Plot a small visualization to make sure things generally look good

```{r}
Nrate_trial <- master_maturity_summed_clean %>% 
  #filter(Year != 2023) %>% 
  filter(Topdress!=1) %>% 
  filter(Stage=="Maturity") %>%
  mutate(Year= as.factor(Year))
  #group_by(Treatment, Nrate_kgha, Year) %>%
  #mutate(N_total_kgha_se = sd(N_total_kgha)/sqrt(3)) %>%
  #summarise(N_total_kgha = mean(N_total_kgha),
  #N_total_kgha_se = mean(N_total_kgha_se))


all_Nuptake_graph <-
ggplot(Nrate_trial, aes(x=Nrate_kgha, y=N_total_kgha, color=Treatment,shape = Year))+
  geom_point(size=2.5)+
  geom_smooth(aes(group = Treatment),method = lm, alpha=0.5)+
  scale_color_manual(values=c("#0072B2","#FFCC66"), name = "Treatment", labels = c("Continuous rice (CR)", "Fallow rice (FR)"))+
  scale_x_continuous(name=expression("N Rate (kg N ha"^{-1}*")"), limits = c(-20, 300), expand = c(0, 0), breaks = seq(0, 300, by = 50))+
  scale_y_continuous(name=expression("N Uptake (kg N ha"^{-1}*")"), limits = c(0, 300), expand = c(0, 0), breaks = seq(0, 300, by = 50))+
  #geom_errorbar(aes(ymin=N_total_kgha-N_total_kgha_se, ymax=N_total_kgha+N_total_kgha_se), width=3,position=position_dodge(0.5))+
  #geom_vline(xintercept = c(41, 50, 78, 84, 121, 136), linetype = "dashed", color = "black") +
  theme_classic()+
  theme(axis.text = element_text(size = 14), axis.title = element_text(size=16))+
  theme(legend.text = element_text(size = 12),legend.title = element_text(size = 14))+
  theme(plot.title = element_text(hjust = 0.5, size = 15))


ggsave(filename = "all_Nuptake_graph.png",  # Include the file extension here
       plot = all_Nuptake_graph,            # Specify the plot
       path = "D:/Academics/UC Davis/School Work/Linquist Lab/Data/R stats/Agronomic paper/Figures",
       dpi = 400,
       height = 30, width = 30, units = "cm")

all_Nuptake_graph

#plots seem to make sense. Lets proceed with more fun things!!!!
```
